\chapter{Data Required}
\label{ch:data-required}

The entire project makes heavy use of synthetic data in form of library dependencies and hypothetical data in the form of chemical formulas as input. As such no human data was required and therefore, no testing was needed. For the input data, several test cases were provided by Dr Mike Gaultois containing desired chemical composition, available reagents as well as real results obtained from theoretical calculations by Dr. Berhard Leube. \\

Other third-party software used, which is not a library, is entirely free and readily available to the public. Microsoft Visual Studio Community was used during development for building the program, as well as NodeJS for running the web-server and Mongo for running the actual database server. Optional but proven very useful for data inspection was MongoDB-Compass, which was utilised to visualise and analyse database entries obtained from our compiled application.\\

A lot of caution has been utilised when selecting 3rd party libraries, and all used have been check and proven to be open-source. No further user attention is required as the program comes packed with all the required library files under a Dynamic-Linked Library (DLL) format.

\section{Ethical Use Of Human Participants}
As with one of our undergraduate projects data from members of the project group does not constitute 3rd party evaluation, and with non-project team members we will be strictly following the CS Department ethical procedure for MEng project 3rd party evaluation. All other data used in the project is well known public data regarding the behaviour of elements. \\